---
title: "Practice Assignment III - Questions"
output:
  html_document:
    df_print: paged
---

## *R*
> Use the information in [this dataset](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/HG7NV7) to answer select questions.

#### Import libraries
```{r}
library(DBI)
library(dbplyr)
library(dplyr)
```

#### Connect to database and list tables
```{r}
conn <- dbConnect(RSQLite::SQLite(), "../airline2.db")
dbListTables(conn)
```

#### Create references ro each table in the database
```{r}
flights <- tbl(conn, "on_time")
planes <- tbl(conn, "planes")
carriers <- tbl(conn, "carriers")
airports <- tbl(conn, "airports")
```

#### Answer questions by running queries against the database

### Which of the following airplanes has the lowest associated average departure delay (excluding Cancelled and Diverted flights)?
```{r}
q1 <- inner_join(flights, planes, by=c("TailNum"="tailnum")) %>%
  filter(Cancelled == 0 & Diverted == 0 & DepDelay > 0) %>%
  group_by(model) %>%
  summarize(avg_delay = mean(DepDelay, na.rm = TRUE)) %>%
  arrange(avg_delay)
q1
```
### Which of the following companies has the highest number of Cancelled flights?
```{r}
q2 <- inner_join(flights, carriers, by=c("UniqueCarrier"="Code")) %>%
  filter(Description %in% c("United Air Lines Inc.", "American Airlines Inc.", "Pinnacle Airlines Inc.", "Delta Air Lines Inc.")) %>%
  group_by(Description) %>%
  summarize(cancellations = sum(Cancelled, na.rm = TRUE)) %>%
  arrange(desc(cancellations))
q2
```

### Which of the following cities has the highest number of inbound flights (excluding cancelled flights)?
```{r}
q3 <- inner_join(flights, airports, by=c("Dest"="iata")) %>%
  filter(Cancelled == 0) %>%
  group_by(city) %>%
  summarize(inbound_flights = n()) %>%
  arrange(desc(inbound_flights))
q3
```

### Which of the following companies has the highest number of cancelled flights, relative to their number of total flights?
```{r}
q4 <- inner_join(flights, carriers, by=c("UniqueCarrier"="Code")) %>%
  filter(Description %in% c("United Air Lines Inc.", "American Airlines Inc.", "Pinnacle Airlines Inc.", "Delta Air Lines Inc.")) %>%
  group_by(Description) %>%
  summarize(numerator = sum(Cancelled, na.rm = TRUE), denominator = n()) %>%
  mutate(ratio = as.double(numerator) / denominator) %>%
  arrange(desc(ratio))
 q4
```

#### Close the database connection
```{r}
dbDisconnect(conn)
```